---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MSETools

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

**MSETools** provides a unified interface to multiple systems estimation (MSE) software. It implements best usage practices and computational speedups. Data from multiple system estimation studies of human trafficking has been reproduced for illustrations and analyses. Analyses from Binette and Steorts (2021) are contained in the "analyses" folder.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("OlivierBinette/MSETools")
```

## Summary


### Implemented models

  - **dga:** The `dga()` function provides an interface to the `dga`
    package of Lum, Johndrow and Ball (2015), here re-implemented using
    Rcpp and extended to allow more flexible priors as well as Bayesian
    fractional posterior distributions. The package implements
    decomposable graphical models with hyper-Dirichlet priors and
    Bayesian model averaging.
  - **LCMCR:** The `lcmcr()` function provides an interface to the
    `LCMCR` package of Manrique-Vallier (2020), which implements the
    latent class model of Manrique-Vallier (2016). By default,
    **MSETools** initializes 500 parallel MCMC chains to provide
    cross-replication stability – this is necessary since the LCMCR
    Gibbs sampler fails to converge in some cases. Convergence
    diagnostics are available through `MSETools::diagnostics()`.
  - **SparseMSE:** The `sparsemse()` function provides an interface to
    the `SparseMSE` package of Chan, Silverman and Vincent (2019), which
    implements a Poisson log-linear approach with stepwise model
    selection and bootstrap confidence intervals. An option has been
    added to parallelize bootstrap replications across available cores.

### Helper functions

  - `estimates()` computes point estimates and confidence intervals for
    a list of models.
  - `batch.estimates()` compute estimates as a slurm job array (for use in a cluster).
  - `summary()` provides a result summary of fitted models.
  - `diagnostics()` provides convergence diagnostics for `lcmcr`
    objects.

### Datasets

See Binette and Steorts (2021) for a description of the datasets reproduced herein.


## Examples

```{r}
library(MSETools)
```

Define a list of models fitted to the UK dataset:

```{r, cache=TRUE}
models = list(lcmcr(UK), sparsemse(UK), dga(UK), independence(UK))
```

Compute estimates:

```{r, cache=TRUE}
estimates(models)
```

Parallelize on a computing cluster:

```{r eval=FALSE}
batch.estimates(models, njobs=4)
```

## Running analyses

Analyses contained in .Rmd files can be knitted on any platform using the included cache. They rely on a computing cluster using SLURM to generate cache files.

## References

- Binette, O. and Steorts, Rebecca C. (2021) On the Reliability of Multiple Systems Estimation for the Quantification of Modern Slavery.
- Lax Chan, Bernard Silverman and Kyle Vincent (2019). SparseMSE: 'Multiple Systems
  Estimation for Sparse Capture Data'. R package version 2.0.1.
  https://CRAN.R-project.org/package=SparseMSE
- James Johndrow, Kristian Lum and Patrick Ball (2021). dga: Capture-Recapture
  Estimation using Bayesian Model Averaging. R package version 2.0.1.
  https://CRAN.R-project.org/package=dga
- Daniel Manrique-Vallier (2020). LCMCR: Bayesian Non-Parametric Latent-Class
  Capture-Recapture. R package version 0.4.11. https://CRAN.R-project.org/package=LCMCR
